High-Performance Language Primitives for Distributed Architectures using Asynchronous Events

### 2012-07-23 doc/asplos2013/micro.tex
section 4 edits

\subsection{Event Latency and Trigger Rates}
\label{subsec:eventmicro}
We implemented two micro-benchmarks for evaluating the performance of
events.  The first micro-benchmark tests the latency of event triggering,
both within a node and between nodes.  Processors are organized
into a ring and each processor in turn creates a user event that is dependent
on the previous processor's event.  The first event in this long chain of
dependent events is then triggered, and the time until the triggering of the
last event in the chain is measured.  The total time is divided by the number
of events in the chain to yield the mean trigger times.  In the single-node case, all events are local to
that node, so no active messages are required.  For all other cases, the
ring uses only a single processor per node so that every trigger requires
the transmission (and reception) of an event trigger active message.

@@

Our second micro-benchmark for events measures the maximum rate at which events can
be triggered by our implementation.  Instead of a single chain, many parallel chains
are created.  Additionally, each step of the chain can consist of parameterized number
of events, each one dependent on all the events from the previous step of the chain.
This parameter is called the {\em fan-in/out factor}.  The events within each step of
a chain are distributed across the nodes.  Recall from Section~\ref{subsec:eventimpl}
that the aggregation of event subscriptions
will limit the number of event trigger active messages to one per node (per event trigger)
even when the fan-in/out factor exceeds the node count.

@@

Figure~\ref{fig:eventthroo} shows the event trigger rates achieved by our implementation
for a variety of node counts and fan-in/out factors.
For small fan-in/out factors, the total rate falls off initially going to two nodes as active
messages become necessary, but starts increasing slightly again at larger node counts.  Higher
fan-in/out factors require more messages and have lower throughput that also increases with
node count.  Although the number of events waiting on each node decreases with an increasing
node code, the minimal scaling indicates that the main bottleneck is in the processing of the
active message that each node must receive rather than the local redistribution of the triggering
notification.
%Note we can further improve performance by increasing
%the number of threads per node for handling active messages (in these experiments there
%is one per node), but in real applications these threads would consume hardware cores and
%take away computational resources from the primary application.  This is only a viable
%option for memory or communication bound programs.

@@

\subsection{Reduction Throughput}
\label{subsec:reducmicro}
To show the benefits of associating reduction instances we designed
a histogram micro-benchmark that requires all nodes to perform an addition reduction to a
physical region located in the globally visible GASNet memory (described in 
Section~\ref{sec:impl}).  Using the reduction interface described in Section~\ref{subsec:phyreg}
reductions are performed in five ways:

@@

We ran two experiments, one corresponding with dense reductions and the other corresponding with sparse
reductions.  In each case a large random source of data is divided across into chunks, which are given to
separate reduction tasks.  There are 8 reduction tasks created for each node, one per processor.  The dense
case uses a histogram with 256K buckets and each reduction task performs 4M reductions.  The results for the
dense experiment are shown in Figure~\ref{fig:reducdense}.  The sparse case use a histogram with 4M buckets,
but only 64K reductions are performed by each task.  Results for the sparse case
be seen in Figure~\ref{fig:reducsparse}.

The dense experiment demonstrates that reduction fold regions perform 
best and scales well with the number of nodes, achieving over a billion reductions
per second in the 16 node case.  List regions also scale well, but perform about an
order of magnitude worse than folding regions in the dense case.  The use of a separate active message for
each reduction operation is another two orders of magnitude worse, demonstrating that data must be
transferred in larger blocks to be efficient on modern interconnect hardware.  The localized instance approach
works well for a single node, but its inherently serial nature doesn't benefit from increasing node counts.  Finally, the only time the latency of performing individual RMA reads and writes isn't disastrous is
on a single node (where the reads and writes are all local).

>>>>

\subsection{Event Latency and Trigger Rates}
\label{subsec:eventmicro}
We implemented two micro-benchmarks for evaluating the performance of
events.  The first micro-benchmark tests the latency of event triggering,
both within a node and between nodes.  Processors are organized
into a ring and each processor in turn creates a user event that is dependent
on the previous processor's event.  The first event in this long chain of
dependent events is then triggered, and the time until the triggering of the
last event in the chain is measured.  The total time is divided by the number
of events in the chain to yield the mean trigger time.  In the single-node case, all events are local to
that node, so no active messages are required.  For all other cases, the
ring uses only a single processor per node so that every trigger requires
the transmission (and reception) of an event trigger active message.

@@

Our second micro-benchmark for events measures the maximum rate at which events can
be triggered by our implementation.  Instead of a single chain, many parallel chains
are created.  Additionally, each step of the chain consists of a parameterized number
of events, each one dependent on every event from the previous step of the chain.
This parameter is called the {\em fan-in/out factor}.  The events within each step of
a chain are distributed across the nodes.  Recall from Section~\ref{subsec:eventimpl}
that the aggregation of event subscriptions
will limit the number of event trigger active messages to one per node (per event trigger)
even when the fan-in/out factor exceeds the node count.

@@

Figure~\ref{fig:eventthroo} shows the event trigger rates achieved by our implementation
for a variety of node counts and fan-in/out factors.
For small fan-in/out factors, the total rate falls off initially going to two nodes as active
messages become necessary, but starts increasing slightly again at larger node counts.  Higher
fan-in/out factors require more messages and have lower throughput that also increases with
node count.  Although the number of events waiting on each node decreases with an increasing
node count, the minimal scaling indicates that the main bottleneck is in the processing of the
active message that each node must receive rather than the local redistribution of the triggering
notification.
%Note we can further improve performance by increasing
%the number of threads per node for handling active messages (in these experiments there
%is one per node), but in real applications these threads would consume hardware cores and
%take away computational resources from the primary application.  This is only a viable
%option for memory or communication bound programs.

@@

\subsection{Reduction Throughput}
\label{subsec:reducmicro}
To show the benefits of reduction instances we designed
a histogram micro-benchmark that requires all nodes to perform an addition reduction to a
physical region located in the globally visible GASNet memory (described in 
Section~\ref{sec:impl}).  Using the reduction interface described in Section~\ref{subsec:phyreg}
reductions are performed in five ways:

@@

We ran two experiments, one corresponding with dense reductions and the other corresponding with sparse
reductions.  In each case a large random source of data is divided into chunks, which are given to
separate reduction tasks.  There are 8 reduction tasks created for each node, one per processor.  The dense
case uses a histogram with 256K buckets and each reduction task performs 4M reductions.  The results for the
dense experiment are shown in Figure~\ref{fig:reducdense}.  The sparse case uses a histogram with 4M buckets,
but only 64K reductions are performed by each task.  Results for the sparse case
be seen in Figure~\ref{fig:reducsparse}.

The dense experiment demonstrates that reduction fold regions perform 
best and scale well with the number of nodes, achieving over a billion reductions
per second in the 16 node case.  List regions also scale well, but perform about an
order of magnitude worse than folding regions in the dense case.  The use of a separate active message for
each reduction operation is another two orders of magnitude worse, demonstrating that data must be
transferred in larger blocks to be efficient on modern interconnect hardware.  The localized instance approach
works well for a single node, but its inherently serial nature doesn't benefit from increasing node counts.  Finally, the only time the latency of performing individual RMA reads and writes isn't disastrous is
on a single node (where the reads and writes are all local).
